### 2011-02-10 paper/hyperNeatMethod.tex

A CPPN is queried once for each link in the ANN phenotype to determine
that link's weight. The inputs to the CPPN are the Cartesian
coordinates of both the source (e.g.\ $x = 2$, $y = 4$) and target
(e.g.\ $x = 3$, $y = 5$) nodes of a link and a constant bias
value. The CPPN takes these five values as inputs and produces two output values. The first output value
determines the weight of the link between the associated input
(source) and hidden layer (target) nodes, and the second output value
determines the weight of the link between the associated hidden
(source) and output (target) layer nodes. All pairwise combinations of
source and target nodes are iteratively passed as inputs to a CPPN to
determine the weight of each ANN link.

@@

\acmFig{SpiderANN.jpg}{1}{ANN Configuration for HyperNEAT Runs. The
  first two columns of each row of the input layer receive information
  about a single leg (the current angle of its two joints). The final
  column provides a sine and cosine wave to enable periodic movements
  and the angle of the center joint. Evolution determines the function
  of the hidden-layer nodes. The nodes in the output layer specify new
  joint angles for each respective joint. The unlabeled node in the
  input and output layer is ignored.}\label{hyperneatLayout}

The population size for HyperNEAT was 9 and runs lasted 20
generations. These numbers are small, but were necessarily constrained
given how much time it took to conduct evolution directly on a real
robot. The remaining parameters were identical to Clune et
al.~\cite{clune2011performance}.

>>>>

A CPPN is queried once for each link in the ANN phenotype to determine
that link's weight~(\figref{hyperneatExplanation.png}). The inputs to the CPPN are the Cartesian
coordinates of both the source (e.g.\ $x = 2$, $y = 4$) and target
(e.g.\ $x = 3$, $y = 5$) nodes of a link and a constant bias
value. The CPPN takes these five values as inputs and produces two output values. The first output value
determines the weight of the link between the associated input
(source) and hidden layer (target) nodes, and the second output value
determines the weight of the link between the associated hidden
(source) and output (target) layer nodes. All pairwise combinations of
source and target nodes are iteratively passed as inputs to a CPPN to
determine the weight of each ANN link.

@@

\acmFig{SpiderANN.jpg}{1}{ANN Configuration for HyperNEAT Runs. The
  first two columns of each row of the input layer receive information
  about a single leg (the current angle of its two joints). The final
  column provides a sine and cosine wave to enable periodic movements
  and the angle of the center joint. Evolution determines the function
  of the hidden-layer nodes. The nodes in the output layer specify new
  joint angles for each respective joint. The unlabeled nodes in the
  input and output layers are ignored.}\label{hyperneatLayout}

As with the parameterized learning methods, three runs of HyperNEAT were performed. The population size for HyperNEAT was 9 and runs lasted 20
generations. These numbers are small, but were necessarily constrained
given how much time it took to conduct evolution directly on a real
robot. The remaining parameters were identical to Clune et
al.~\cite{clune2011performance}.

### 2011-02-10 paper/learningMethods.tex
minor
hyperneat edits

By a \emph{parametrized gait}, we mean a gait produced by a
parametrized function $g(t; \vec{\theta})$. Fixing the parameter
$\vec{\theta}$ yields a deterministic motion function over time.  We
tried several parametrizations on the robot and, upon obtaining
reasonable early success, settled on one particular parametrization, which we call  \emph{SineModel5}. Its root pattern is a sine wave and it has five parameters (Table~\ref{tab:params}) .

@@

\edit{called outer multiplier, but text says we multiply by inner. same for left}

Intuitively, SineModel5 starts with 8 identical sine waves of
amplitude $\amp$ and period $\tau$, multiplies the waves for all inner
motors by $m_O$, multiplies the waves for all front motors by $m_F$,
and multiplies the waves for all left motors by $m_R$.  To obtain the
actual motor position commands, these waves are offset by
appropriate fixed constants so that the base position (when the sine
waves are at 0) is approximately a crouch (the position shown in
\figref{robot_close.jpg}).  In this motion model, the ninth
(center) motor is kept at a neutral position.  Thus, the commanded
position for all motors, as a vector function of time, is:

@@

\noindent Here the fixed constants that determine the crouched
position are $C_I = 800$ for inner motors, $C_O = 40$ for outer
motors, and $C_C = 512$ for the center hip motor.  The nine functions
correspond to the nine motors in the same order as listed in
\figref{topdown.png}.

\edit{what subscripts?}

@@

\item \emph{Linear regression}: To initialize, this method chooses and
  evaluates five random parameter vectors. It then fits a linear model
  from parameter vector to fitness. In a loop, the method chooses and
  evaluates a new parameter vector generated by taking a fixed-size
  step in the direction of the gradient for each parameter, and fits a
  new linear model to all vectors evaluated so far, choosing the model
  to minimize the sum of squared errors. Step size is the same as in
  \emph{N-dimensional policy gradient ascent}.

>>>>

By a \emph{parametrized gait}, we mean a gait produced by a
parametrized function $g(t; \vec{\theta})$. Fixing the parameter
$\vec{\theta}$ yields a deterministic motion function over time.  We
tried several parametrizations on the robot and, upon obtaining
reasonable early success, settled on one particular parametrization, which we call  \emph{SineModel5}. Its root pattern is a sine wave and it has five parameters (\tabref{params}).

@@

Intuitively, SineModel5 starts with 8 identical sine waves of
amplitude $\amp$ and period $\tau$, multiplies the waves for all outer
motors by $m_O$, multiplies the waves for all front motors by $m_F$,
and multiplies the waves for all right motors by $m_R$.  To obtain the
actual motor position commands, these waves are offset by
appropriate fixed constants so that the base position (when the sine
waves are at 0) is approximately a crouch (the position shown in
\figref{robot_close.jpg}).  In this motion model, the ninth
(center) motor is kept at a neutral position.  Thus, the commanded
position for all motors, as a vector function of time, is:

@@

\noindent Here the fixed constants that determine the crouched
position are $C_I = 800$ for inner motors, $C_O = 40$ for outer
motors, and $C_C = 512$ for the center hip motor.  The nine functions (in order)
correspond to the nine motors listed in
\figref{topdown.png}.

@@

\item \emph{Linear regression}: To initialize, this method chooses and
  evaluates five random parameter vectors. It then fits a linear model
  from parameter vector to fitness. In a loop, the method chooses and
  evaluates a new parameter vector generated by taking a fixed-size
  step in the direction of the gradient for each parameter, and fits a
  new linear model to all vectors evaluated so far, choosing the model
  to minimize the sum of squared errors. The step size is the same as in
  \emph{N-dimensional policy gradient ascent}.

Each of the seven learning methods was run on 3 different initial
parameter vectors ($\vec{\theta}_A$,
$\vec{\theta}_B$, and $\vec{\theta}_C$) in order to fairly compare the algorithms. Three runs were performed per learning method. Runs continued until the performance plateaued, which we defined as when there was no improvement during the last third of
a run.
